### 2017-04-05 writing/paper/sections/conclfut.tex
Make Changes suggest by Michael
All comments from Henning. See TODOs

If we consider a similar system for dominating patterns of length \(4\) and
mesh patterns of length \(2\), it can be seen that the number of cases required
to establish rules increases to a number that is infeasible to compute manually.
For an extension of the First Dominating rule alone, we would have to consider
placement of points in any pair of unshaded regions. The fact that the rules
established do not completely cover the coincidences with a dominating pattern
of length \(3\) shows that this is a difficult task.

@@

   Consider a permutation containing \(m_1\),
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{}{1,2}{0/1,0/2,1/1,1/2,2/0}
         \draw (1.5,0.5) node {\(Y\)};
         \draw (2.5,1.5) node {\(X\)};
     \end{tikzpicture}
   \end{equation*}
   If the regions corresponding to both \(X\) and \(Y\) are empty then we have
   an occurrence of \(m_2\).
   Consider if the region corresponding to \(X\) is non-empty, we can then choose
   the lowest valued point in this region
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{1,3}{1,3,2}{0/1,0/2,0/3,1/1,1/2,1/3,2/0,2/1,3/0,3/1}
         \draw (1.5,0.5) node {\(Y\)};
     \end{tikzpicture}
   \end{equation*}
   If the region corresponding to \(Y\) is empty then we have an occurrence of
   \(m_2\) with the indicated points.
   Now if the region corresponding to \(Y\) is non-empty, we can choose the
   rightmost point in this region.
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{2,4}{2,1,4,3}{0/2,0/3,0/4,1/2,1/3,1/4,2/0,2/1,2/2,2/3,2/4,3/0,3/1,3/2,4/0,4/1,4/2}
     \end{tikzpicture}
   \end{equation*}
   And now the two indicated points form an occurrence of \(m_2\).
   We have therefore shown that any occurrence of \(m_1\) is an occurrence of
   \(m_2\) and we can easily show the converse by the same reasoning, so \(m_1\)
   and \(m_2\) are coincident.
   This is captured by an extension of the Third Dominating rule where we allow
   multiple steps of adding points before we check for subpattern containment.
 \end{example}

It would be interesting to consider a systematic explanation of Wilf-equivalences among classes
where \(\perm{2,3,1}\) is the dominating pattern, possibly using the  construction presented in
\cite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

>>>>

If we consider a similar approach to dominating patterns of length \(4\) and
mesh patterns of length \(2\), it can be seen that the number of cases required
to establish rules increases to a number that is infeasible to compute manually.
For an extension of the First Dominating rule alone, we would have to consider
placement of points in any pair of unshaded regions. The fact that the rules
established do not completely cover the coincidences with a dominating pattern
of length \(3\) shows that this is a difficult task.

@@

   Consider a permutation containing \(m_1\),
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{}{1,2}{0/1,0/2,1/1,1/2,2/0}
         \draw (1.5,0.5) node {\(Y\)};
         \draw (2.5,1.5) node {\(X\)};
     \end{tikzpicture}
   \end{equation*}
   If the regions corresponding to both \(X\) and \(Y\) are empty then we have
   an occurrence of \(m_2\).
   If the region corresponding to \(X\) is non-empty, we can then choose
   the lowest valued point in this region
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{1,3}{1,3,2}{0/1,0/2,0/3,1/1,1/2,1/3,2/0,2/1,3/0,3/1}
         \draw (1.5,0.5) node {\(Y\)};
     \end{tikzpicture}
   \end{equation*}
   If the region corresponding to \(Y\) is empty then we have an occurrence of
   \(m_2\) with the indicated points.
   Now if the region corresponding to \(Y\) is non-empty, we can choose the
   rightmost point in this region.
   \begin{equation*}
     \begin{tikzpicture}[scale=0.5]
         \modpattern[5]{2,4}{2,1,4,3}{0/2,0/3,0/4,1/2,1/3,1/4,2/0,2/1,2/2,2/3,2/4,3/0,3/1,3/2,4/0,4/1,4/2}
     \end{tikzpicture}
   \end{equation*}
   And now the two indicated points form an occurrence of \(m_2\).
   We have therefore shown that any occurrence of \(m_1\) leads to an occurrence of
   \(m_2\) and we can easily show the converse by the same reasoning, so \(m_1\)
   and \(m_2\) are coincident.
   This is captured by an extension of the Third Dominating rule where we allow
   multiple steps of adding points before we check for subpattern containment.
 \end{example}

It would be interesting to consider a systematic explanation of Wilf-equivalences among classes
where \(\perm{3,2,1}\) is the dominating pattern, possibly using the  construction presented in
Bean, Claesson and Ulfarsson~\cite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

### 2017-04-05 writing/paper/sections/wilf.tex
All comments from Henning. See TODOs

Wilf-equivalence is an important aspect to study in the field of permutation
patterns. Two patterns \(\pi\) and \(\sigma\) are said to be Wilf-equivalent
if their avoidance sets have the same size at each length. More formally:

@@

Coincident patterns are trivially Wilf-equivalent, if \(\av[k]{R} = \av[k]{S}\)
then trivially \(\size{\av[k]{R}} = \size{\av[k]{S}}\). Coincidence is
therefore a stronger equivalence condition than Wilf-equivalence.

@@

Since we are always considering Wilf-equivalences in the set \(\av{S}\) we
must only use these symmetries when they preserve the dominating pattern, if we
were to allow other symmetries, then the equivalences calculated in the
previous section do not necessarily hold.

Throughout this section we will consider Wilf-equivalences of patterns
whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\). We will
use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
be the usual Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\).
The fact that \(C(x)\) is teh generating function for \(\mathcal{C}\)
can be seen by structural decomposition around the maximum, as
shown in \FigureRef{fig:decompmax}.

@@

\begin{equation*}
    \scriptvar = \varepsilon \sqcup
\decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
If the upper-right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\), so we can use any avoider of \(\perm{2,3,1}\). The lower-left
section however can create occurrences
of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}\\
    \var(x) &= \frac{C(x)}{1+xC(x)}
\end{align*}

Calculating coefficients given by this generating function gives the Fine
numbers.
\begin{equation*}
    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc
\tag{\href{https://oeis.org/A000957}{OEIS: A000957}}
\end{equation*}

\nextvar[\varfine]
It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident. Consider the decomposition of a non-empty avoider of \(q_1\) in
\(\av{\perm{2,3,1}}\) around the maximum:
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup
\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(q_1\) starts with its maximum, by not allowing the lower-left section of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever
happening.

Consider \(q_2 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
means that a permutation does not end with its maximum. We can perform a
similar
decomposition as before to get
\begin{equation*}
    \mathcal{\var}_2 =\varepsilon \sqcup
\decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}

Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
pattern are permutations that do not start with their minimum. In this case we
perform the decomposition around the leftmost element
\begin{equation*}
    \mathcal{\var}_3 =\varepsilon \sqcup
\decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
All of these classes have the same generating function, namely
\begin{equation}
    \var(x) = 1 + xC(x)(C(x)-1). \label{eqn:maxlgen}
\end{equation}
The coefficients of this generating function are
\begin{equation*}
    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
\tag{\href{https://oeis.org/A000245}{OEIS: A000245} with offset \(1\)}
\end{equation*}

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Any avoider of \(\perm{2,3,1}\) that
ends in its minimum must be a decreasing sequence
Therefore this particular class has equation
\begin{equation*}
    \var(0) = 1, \var(x) = C(x)-1
\end{equation*}
Computing these values gives
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
\tag{\href{https://oeis.org/A141364}{OEIS: A141364}}
\end{equation*}

\section{Wilf-classes with patterns of length 2}
By the use of set equivalences established in
\SectionRef{sec:coincs} we know that there are at most 95 Wilf-equivalence classes.

@@

Computing avoiders up to length 10 suggests that there are at least 23 Wilf-classes,
of which 13 are non-trivial, therefore there are Wilf-equivalences that are not
explained by coincidences or symmetry.

@@

\begin{definition} A \emph{Set Partition} of \(\nrange{n}\) is a set of non-empty subsets of \(\nrange{n}\)
such that \(\nrange{n}\) is a \emph{disjoint} union of these subsets.
\end{definition}

@@

We will call the least element in each block the \emph{block-bottom}, and note that all permutations
in \(\av{\perm{2,3,1}}\) also avoid this mesh pattern.

@@

First we prove the Wilf-equivalence between \(m_1\) and \(m_2\) shown in
\eqref{eq:singleud}. The easiest way to
show that these are equinumerous is to consider the containers as set partitions.

@@

Looking at avoiders of \(\perm{2,3,1}\) and \(m_4\) we
can perform a similar decomposition around the maximum to get
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\scriptvar_2}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
Any occurrence of \(m_4\) can never occur in the
top right region. It could only occur between the maximum and the first point
in the region, if and only if this first point is the lowest valued element in
this region.
Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
enumeration, \(\scriptvar_1\) and \(\scriptvar_2\) must also have
the same enumeration and are therefore Wilf-equivalent.

Now we must consolidate these two subclasses. In order to do this we must
consider the decomposition around the leftmost point of a permutation in
\(\av{\perm{2,3,1},m_1}\). We have the following
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompleft{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
have the same enumeration, and therefore all four patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\) with generating function satisfying
\begin{equation*}
    \var(x) = 1 + x\var(x)\varmaxl(x)
\end{equation*}
Where \(\varmaxl(x)\) is the generating function given in \EquationRef{eqn:maxlgen}.
This can be enumerated to give the sequence
\begin{equation*}
    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc \tag{\href{https://oeis.org/A035929}{OEIS: A035929} offset 1}
\end{equation*}

@@

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \text{ and }
    m_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
\end{equation*}
Let \(\scriptvar_1\) be the set of avoiders of \(m_1\) in \(\av{231}\). By structural decomposition
around the leftmost point we have
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptvar_1}{}{}{\scriptvar^\prime_1}
\end{equation*}
Where \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
and \(\textpattern{}{1}{0/0,0/1,1/0}\).
Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\). It can
once again be decomposed around the leftmost point
\begin{equation*}
    \scriptvar^\prime_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_1\setminus\varepsilon}{}{}{\scriptvar^\prime_1}
\end{equation*}
This is a complete decomposition of avoiders of \(m_1\). Now we
look at an avoider of \(m_2\), this time decomposition is around the leftmost point
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptvar^\prime_2}{}{}{\scriptvar_2}
\end{equation*}

Where \(\scriptvar^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\).
Again we use the same method of decomposition of a permutation in \(\scriptvar^\prime_2\)
\begin{equation*}
    \scriptvar^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptvar^\prime_2}{}{}{\scriptstyle \scriptvar_2\setminus\varepsilon}
\end{equation*}
This gives us a generating function \(\var(x)\) satisfying
\begin{align}
    \var(x) &= 1 + x\var(x)\var^\prime(x) \label{eqn:Pgen}\\
    \var^\prime(x) &= 1 + x(\var(x)-1)\var^\prime(x)\label{eqn:PprimeGen}
\end{align}
Solving \EquationRef{eqn:PprimeGen} for \(\var^\prime(x)\) and substituting into
\EquationRef{eqn:Pgen} gives us that the generating function for
\(\var(x)\) satisfies
\begin{equation}
    \var(x) = x\var^2(x) - x(\var(x) - 1) + 1
\end{equation}
Evaluating \(\var(x)\) gives us the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc \tag{\href{https://oeis.org/A001006}{OEIS: A001006 with offset 1}}
\end{equation*}
This is an offset of the Motzkin numbers.

>>>>

Wilf-equivalence is an important aspect to study in the field of permutation
patterns.

@@

Coincident patterns are trivially Wilf-equivalent: if \(\av[k]{R} = \av[k]{S}\)
then trivially \(\size{\av[k]{R}} = \size{\av[k]{S}}\). Coincidence is
therefore a stronger equivalence condition than Wilf-equivalence.

@@

Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
must only use these symmetries when they preserve the dominating pattern, if we
were to allow other symmetries, then the equivalences calculated in the
previous section do not necessarily hold.

Throughout this section we will consider Wilf-equivalences of patterns
whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\). We will
use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
be the usual Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\).
The fact that \(C(x)\) is the generating function for \(\mathcal{C}\)
can be seen by structural decomposition around the maximum, as
shown in \FigureRef{fig:decompmax}.

@@

\begin{equation*}
    \scriptvar = \varepsilon \sqcup
\decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
If the upper-right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\), so we can use any avoider of \(\perm{2,3,1}\). The lower-left
section however can create occurrences
of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}\\
    \var(x) &= \frac{C(x)}{1+xC(x)}
\end{align*} % TODO Do we need both representations?

\noindent
Calculating coefficients given by this generating function gives the Fine
numbers.
\begin{equation*}
    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc
\tag{\href{https://oeis.org/A000957}{OEIS: A000957}}
\end{equation*}

\nextvar[\varfine]
It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident. Consider the decomposition of a non-empty avoider of \(q_1\) in
\(\av{\perm{2,3,1}}\) around the maximum:
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup
\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(q_1\) starts with its maximum, and by not allowing the lower-left section of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever
happening.

Consider \(q_2 = \textpattern{}{1}{0/1,1/0,1/1}\). Avoiding this pattern
means that a permutation does not end with its maximum. We can perform a
similar
decomposition as before to get
\begin{equation*}
    \mathcal{\var}_2 =\varepsilon \sqcup
\decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}

% TODO Isn't q_3 a symmetry of q_2? So we don't need to repeat the work
Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\). The avoiders of this
pattern are permutations that do not start with their minimum. In this case we
perform the decomposition around the leftmost element
\begin{equation*}
    \mathcal{\var}_3 =\varepsilon \sqcup
\decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
All of these classes have the same generating function, namely
\begin{equation}
    \var(x) = 1 + xC(x)(C(x)-1). \label{eqn:maxlgen}
\end{equation}
The coefficients of this generating function are
\begin{equation*}
    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
\tag{\href{https://oeis.org/A000245}{OEIS: A000245} with offset \(1\)}
\end{equation*}

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Any avoider of \(\perm{2,3,1}\) that
ends in its minimum must be a decreasing sequence.
Therefore this particular class has equation
\begin{equation*}
    \var(0) = 1, \var(x) = C(x)-1
\end{equation*}
Computing these values gives
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
\tag{\href{https://oeis.org/A141364}{OEIS: A141364}}
\end{equation*}

\section{Wilf-classes with patterns of length 2}
By the use of coincidence classes established in
\SectionRef{sec:coincs} we know that there are at most 95 Wilf-equivalence classes.

@@

Computing avoiders up to length 10 suggests that there are at least 23 Wilf-classes,
of which 13 are non-trivial. Therefore there are Wilf-equivalences that are not
explained by coincidences or symmetry.

@@

\begin{definition} A \emph{set partition} of \(\nrange{n}\) is a set of non-empty subsets of \(\nrange{n}\)
such that \(\nrange{n}\) is a \emph{disjoint} union of these subsets.
\end{definition}

@@

We will call the least element in each block the \emph{block-bottom}, and note that all permutations
in \(\av{\perm{2,3,1}}\) also avoid the mesh pattern \(q\).

@@

First we prove the Wilf-equivalence between \(m_1\) and \(m_2\) shown in
\eqref{eq:singleud}, by considering the containers as set partitions consider
the containers as set partitions.

@@

Looking at avoiders of \(\perm{2,3,1}\) and \(m_4\) we
can perform a similar decomposition around the maximum to get
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\scriptvar_2}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
An occurrence of \(m_4\) can never occur in the
top right region. It could only occur between the maximum and the first point
in the region, if and only if this first point is the lowest valued element in
this region.
Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
enumeration, \(\scriptvar_1\) and \(\scriptvar_2\) must also have
the same enumeration and are therefore Wilf-equivalent.

Now we must consolidate these two subclasses. In order to do this we must
consider the decomposition around the leftmost point of a permutation in
\(\av{\perm{2,3,1},m_1}\). We have the following
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompleft{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
have the same enumeration, and therefore all four patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\) with generating function satisfying
\begin{equation*}
    \var(x) = 1 + x\var(x)\varmaxl(x)
\end{equation*}
where \(\varmaxl(x)\) is the generating function given in \EquationRef{eqn:maxlgen}.
This can be enumerated to give the sequence
\begin{equation*}
    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc \tag{\href{https://oeis.org/A035929}{OEIS: A035929} offset 1}
\end{equation*}

@@

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \text{ and }
    m_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
\end{equation*}
Let \(\scriptvar_1\) be the set of avoiders of \(m_1\) in \(\av{231}\). By structural decomposition
around the leftmost point we have
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptvar_1}{}{}{\scriptvar^\prime_1}
\end{equation*}

\noindent
Here \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
and \(\textpattern{}{1}{0/0,0/1,1/0}\).
Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\). It can
once again be decomposed around the leftmost point
\begin{equation*}
    \scriptvar^\prime_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_1\setminus\varepsilon}{}{}{\scriptvar^\prime_1}
\end{equation*}
This is a complete decomposition of avoiders of \(m_1\). Now we
look at an avoider of \(m_2\), this time decomposition is around the leftmost point
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptvar^\prime_2}{}{}{\scriptvar_2}
\end{equation*}

Where \(\scriptvar^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\).
Again we use the same method of decomposition of a permutation in \(\scriptvar^\prime_2\)
\begin{equation*}
    \scriptvar^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptvar^\prime_2}{}{}{\scriptstyle \scriptvar_2\setminus\varepsilon}
\end{equation*}
This gives us a generating function \(\var(x)\) satisfying
\begin{align}
    \var(x) &= 1 + x\var(x)\var^\prime(x) \label{eqn:Pgen}\\
    \var^\prime(x) &= 1 + x(\var(x)-1)\var^\prime(x)\label{eqn:PprimeGen}
\end{align}
Solving \EquationRef{eqn:PprimeGen} for \(\var^\prime(x)\) and substituting into
\EquationRef{eqn:Pgen} gives us that the generating function for
\(\var(x)\) satisfies
\begin{equation}
    \var(x) = x\var^2(x) - x(\var(x) - 1) + 1
\end{equation}
Evaluating \(\var(x)\) gives us the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc \tag{\href{https://oeis.org/A001006}{OEIS: A001006 with offset 1}}
\end{equation*}
Which is an offset of the Motzkin numbers.
